Reproducible research project week 2
=========
# Code for reading in the dataset and/or processing the data

## Reading file
```{r}
data <- read.csv("activity.csv")
str(data)
summary(data)
head(data)
```

## Transforming column date to Date type
```{r}
data$date <- as.Date(as.character(data$date))
```

# Histogram of the total number of steps taken each day
```{r}
stepsDay <- with(data,tapply(steps,date,sum,na.rm=TRUE))
hist(stepsDay, main = "Histogram activity",xlab = "Steps per Day", col = "red")
```

# Mean and median number of steps taken each day
```{r}
mean(stepsDay)
median(stepsDay)
```

# Time series plot of the average number of steps taken
```{r}
meanSteps <- with(data,tapply(steps,interval,mean,na.rm=TRUE))
plot(meanSteps,main = "Daily average number of steps",type = "l",col="blue",
     ylab = "Average number of steps",xlab = "5 min interval")
```

# The 5-minute interval that, on average, contains the maximum number of steps
```{r}
maxInterval <- data$interval[grep(max(meanSteps),meanSteps)]
maxInterval
```

# Code to describe and show a strategy for imputing missing data

We will create a function to fill the NA values with the mean value for each 5 minute interval
```{r}
library(dplyr)
meanInt <- function(x){replace(x, is.na(x), mean(x, na.rm = TRUE))}
newData <- data %>% 
  group_by(interval) %>%
  mutate(steps = meanInt(steps))
head(newData)
```

# Histogram of the total number of steps taken each day after missing values are imputed
```{r}
stepsDayNew <- with(newData,tapply(steps,date,sum,na.rm=TRUE))
hist(stepsDayNew, main = "Histogram activity",xlab = "Steps per Day", col = "red")
```

## New Mean and Median steps per day
```{r}
mean(stepsDayNew)
median(stepsDayNew)
```

### Do these values differ from the estimates from the first part of the assignment? 
YES  

### What is the impact of imputing missing data on the estimates of the total daily number of steps?  
The new histogram shows a normal distribution, new mean and new median have the same value

### Are there differences in activity patterns between weekdays and weekends?
```{r}
newData$weekday <- weekdays(data$date)
newData$weekcat <- ifelse(newData$weekday %in% c("Saturday","Sunday"),"Weekend","Weekday")
```

# Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
stepsMean <- aggregate(newData$steps,by=list(newData$interval,newData$weekcat),mean)
colnames(stepsMean) <- c("Interval","Weekcat","Stepsmean")
str(stepsMean)
library(ggplot2)
ggplot(stepsMean,aes(Interval,Stepsmean,color=Weekcat)) +
  geom_line() +
  facet_grid(Weekcat~.) + 
  labs(y="Average number of steps", x="5-minute interval") + 
  theme(legend.position = "none")
```
